---
title: "Gov 50 Final Project"
author: Hanna Johnson
description: "Hanna Johnson's final project for Gov50 on the relationship between countries with gender ministries and violence against women"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction
<!-- an introduction where you introduce the research question and hypothesis and briefly describe why it is interesting;-->
<!-- An introduction section introducing your research question (a paragraph or two) -->
I am interested in exploring whether or not the existence of an office dedicated to the advancement of women's rights within the government is correlated with a lower percentage of violence against women.  For the purpose of this study, I examined agencies that are Cabinet-level or  equivalencies, and will refer to these as "gender ministries," which broadly encompasses multiple options, including gender ministries, equality ministries, and family ministries, among others. I will also be using country's self-reported rates of violence against women as well as a legal data analysis of their laws on whether or not they harm women.  These results could provide helpful metrics for a number of important questions, including the general effectiveness of such offices and positions.  More specifically, it could demonstrate whether there are good mechanisms for reporting within the government, if there are laws that protect women that may have resulted from the existence of such offices, and if they have  reduced gender discrimination, measured either through less violence, or less stigmatizing beliefs about women.  However, due to the nature of this question, it may be difficult to parse through these possibilities.  I affirm that this is still an essential topic for research, despite these difficulties, and if gender ministries are associated with a reduced outcome of violence against women, this makes an important case for their existence in other countries, too. 

I hypothesize that the existence of gender ministries does reduce violence against women and predict that statistics will be lower in countries with these departments, especially if the leader of this department is a woman, as it is well-established that when someone takes office, they are more likely to advocate for the interests of their own group.  In other words, a woman in office, especially a gender-focused office, the office is likely to accomplish something tangible due to the impacts of representation, exposure, and substantive representation.  However, I also acknowledge that the statistics could also be higher, due to an increased accessibility and greater feeling of safety to report violence against women when it occurs. Patterns that would support my hypothesis are if countries with gender ministries reliably had lower rates of violence against women.  Patterns that would not support my hypothesis could include if there appeared to be no statistically significant relationship, or if numbers were significantly higher, although that could also be due to an increased number of reports due to more trust in the system and more protection, not necessarily a higher rate of violence.


## Data
<!-- a data section that briefly describes the data source, describes how the key dependent and independent variables are measured (e.g., a survey, statistical model, or expert coding), and also produces a plot that summarizes the dependent variable;  -->
<!-- for the data section, you should note if your research design is cross-sectional (most projects will be of this type) or one of the other designs we discussed (randomized experiment, before-and-after, differences-in-differences). -->
<!-- A section describing your data, including the type of research design (e.g. experimental, cross-sectional, longitudinal, etc.) and a plot that summarizes the dependent variable (the length here may depend on how complicated your data is, but most data sections will be a couple of paragraphs) -->
I began collecting data by locating the Violence Against Women dataset from the OECD to gather data on (insert country amount) on their lifetime prevalence of violence against women as well as whether or not they have laws that actively harm women.  Prevalence of violence against women is measured by a percent between zero and one hundred, with zero being no violence and one hundred being total violence, and is self-reported by the government.  Laws that actively harm women are measured between zero and one, with zero indicating no existing laws that actively harm women, and one indicating the presence of laws that clearly harm women, and is measured based on the current country's laws and is not self-reported.  Most countries fall somewhere in between the extremes on most counts.  I used rates of violence against women as the dependent variable and physically controlled for the presence of laws that actively harm women, as I believed that could have otherwise influenced my results.  The data source only provided data for one year, 2019, so I will not be conducting any sort of time review in this project.

Next, to establish whether or not a country has a gender ministry, I manually searched for details on (insert country amount) to align with the data from the OECD data source. I measured between zero and one, with zero indicating no gender ministry whatsoever and one indicating some sort of gender ministry.  There were no numbers in between these two extremes, as I combined any sort of gender ministry as a positive number.  In future research, I may consider splitting it further, to distinguish between equality ministries, family ministries, and others.  However, for the purposes of this study, I was unsure how to categorize them in terms of effectiveness, and since all ministries included focus on gender to some high degree, chose to categorize them the same.  Additionally, there are so many different names for the ministries, but they perform very similar roles, and I was not sure how to conduct a unbiased review of these.  Thus, the presence (or lack thereof) of a gender ministry was my independent variable. 

Finally, my research design is cross-sectional, since I will be using reported rates of violence against women and an analysis of existing laws from one point in time.  Additionally, I did not interfere, but rather observed the existing data without any manipulation to deduce if there was any correlation between the existence of gender ministries and the rates of violence against women, when physically controlling for the country's laws and whether or not they specifically harm women.  Below is a plot (Figure 1) which summarizes the dependent variable.

https://data.oecd.org/inequality/violence-against-women.htm  





```{r include=FALSE}
library(tidyverse)

vaw <- read.csv("OECD_VAW.csv")
vaw

laws <- read.csv("LAWS.csv")
laws

prevalence <- read.csv("PREVALENCE.csv")
prevalence 

```
## Results
<!-- a visualization of the data in question that speaks to your research question -->
<!-- a presentation (as a table or graph) of a regression model assessing your question along with a plain-English interpretation -->
<!-- a results section that contains a scatterplot, barplot, or boxplot of the main relationship of interest and output for the main regression of interest;  -->
<!-- For the results section, you should interpret (in plain English) the main coefficient of interest in your regression. You should also comment on the statistical significance of the estimated coefficient and whether or not you believe the coefficient to represent a causal effect. -->
<!-- A results section, including a presentation of your regression results (either a plot or a table), a separate plot of the main relationship of interest (e.g. a scatterplot of the outcome and the main independent variable), and as well as a summary and discussion of results in substantive context (a paragraph or two).  -->

<!-- You should include at least one multiple regression (i.e. regression with multiple explanatory/independent variables) in your analysis, but note that you're not graded on the results themselves (like whether you see any statistical significance). -->

<!-- You should also interpret (in plain English) the main coefficient of interest in your regression, comment on the statistical significance of the estimated coefficient, and  say whether or not you believe the coefficient to represent a causal effect. -->
```{r include=FALSE}
laws$LOCATION[2] <- "AUT"
```

```{r include=FALSE}
no_ministry_countries <- c("HUN","ISL","IRL","USA","ALB","ARM","BLR","CHN","HRV","GEO","HKG","KAZ","LVA","MDA","PHL","ROU","RUS","SAU","SGP","SDN","THA","UKR","ARE","URY","BHR","BTN","BOL","COG","CUB","ECU","SLV","ERI","GAB","GTM","GNB","IRQ","KGZ","LAO","MNG","MMR","OMN","PNG","QAT","SWZ","SYR","TJK","TLS","TGO","TKM","UZB","YEM","BRN","MLT","TWN","PAK","KWT", "SRB")

laws_2 <- laws |>
  mutate(Gender_Ministry = "Gender Ministry") |>
  mutate(Gender_Ministry = if_else(LOCATION %in% no_ministry_countries, "No Gender Ministry", Gender_Ministry))
laws_2
```

```{r echo=FALSE}
laws_plot <- ggplot(data = laws_2) +
geom_histogram(mapping = aes(x = Value, y = ..prop..), stat = "count", binwidth = 0.1) +
  # xlim(0, 1) +
  scale_x_continuous(breaks=c(0, 0.25, 0.5, 0.75, 1),
        labels=c("0", "0.25", "0.5","0.75","1")) +
  labs(x = "Presence of Laws that Harm Women",
       y = "Percentage of Countries with a Gender Ministry")
laws_plot + facet_wrap(~Gender_Ministry)


```


This visualization demonstrates that for countries with a gender ministry, there are fewer countries with harmful laws about violence against women, where the most harmful laws are 1 and the least harmful laws are 0.  It is notable that gender ministry is an expansive category here, including countries with a distinct Ministry of Gender, but also countries with a combined ministry, such as Ministry of Family and Gender, or Ministry of Social Issues with gender as a primary focus.  Countries without a gender ministry are countries that lack all of these ministry options.  Thus, in each category of laws, in categories which mean more harmful laws, including 1 and 0.75 are more highly represented in the countries without a gender ministry or an office tangential to a gender ministry.  For countries with fewer harmful laws, including 0.5 and 0.25, countries with a gender ministry or an office adjacent to it are more highly represented.

```{r include=FALSE}
no_ministry_countries <- c("HUN","ISL","IRL","USA","ALB","ARM","BLR","CHN","HRV","GEO","HKG","KAZ","LVA","MDA","PHL","ROU","RUS","SAU","SGP","SDN","THA","UKR","ARE","URY","BHR","BTN","BOL","COG","CUB","ECU","SLV","ERI","GAB","GTM","GNB","IRQ","KGZ","LAO","MNG","MMR","OMN","PNG","QAT","SWZ","SYR","TJK","TLS","TGO","TKM","UZB","YEM","BRN","MLT","TWN","PAK","KWT", "SRB")

prevalence_2 <- prevalence |>
  mutate(Gender_Ministry = "Gender Ministry") |>
  mutate(Gender_Ministry = if_else(LOCATION %in% no_ministry_countries, "No Gender Ministry", Gender_Ministry))
prevalence_2
```

```{r echo=FALSE}
mean_prevalence_ministry <- prevalence_2 |>
  filter(Gender_Ministry == "Gender Ministry") |>
  summarize(mean_ministry = mean(Value, na.rm = TRUE)) |>
  pull()

mean_prevalence_no_ministry <- prevalence_2 |>
  filter(Gender_Ministry == "No Gender Ministry") |>
  summarize(mean_no_ministry = mean(Value, na.rm = TRUE)) |>
  pull()

prevalence_plot <- ggplot(data = prevalence_2, aes(x = Value, fill = Gender_Ministry)) +
  geom_density(alpha = 0.5) +
  geom_vline(aes(xintercept = mean_prevalence_ministry), color = "red") +
  geom_vline(aes(xintercept = mean_prevalence_no_ministry), color = "blue") +
# geom_histogram(mapping = aes(x = Value), stat = "count") +
  # xlim(0, 100) +
  # scale_x_continuous(breaks=c(0, 25, 50, 75, 100),
  #       labels=c("0", "25", "50","75","100")) +
  labs(x = "Reported Prevalence of Violence Against Women",
       y = "Density",
       fill = "")
prevalence_plot
# prevalence_plot + facet_wrap(~Gender_Ministry) 
```
This was my first attempt at trying to answer my research question on whether or not having a Ministry of Gender is correlated with better outcomes for women, particularly on violence against women.  This visualization demonstrates that the rates of violence against women is similar for countries with and without a Gender Ministry, with higher reported rates of violence for countries with a Gender Ministry, although it is currently unclear whether or not that violence is statistically significant. The prevalence is measured in a percent of women who both experience and report violence against women, ranging from 0 to 100, although no countries reported a rate of 100.  However, it is notable to mention that over 30 countries included in the original laws data, displayed in my first visualization, did not submit reports of violence against women, many of which are countries that do not have a gender ministry.  Thus, these results could be skewed by the fact that many countries, including those without a Gender Ministry, did not submit and may have interfered with the data.  Moreover, this also relies on reported violence against women, which is infamously extremely under-reported, so it is possible that higher reported prevalence of violence against women could actually be positive, since it could mean that women felt safe enough to report violence, which could have been facilitated by the presence of a Gender Ministry.
```{r}
prevalence_and_laws <- prevalence |>
  pivot_wider(names_from = SUBJECT, values_from = Value)
prevalence_and_laws

laws_2 <- laws_2 |>
  pivot_wider(names_from = SUBJECT, values_from = Value)
laws_2

prevalence_and_laws_2 <- merge(laws_2, prevalence_and_laws, by = c("LOCATION", "TIME"))

prevalence_and_laws_2 <- prevalence_and_laws_2 |>
  select(c(LOCATION, TIME, Gender_Ministry, LAWDOMVIOL, PREVVIOLLIFETIME))
prevalence_and_laws_2
```
```{r}
prevalence_and_laws_3 <- prevalence_and_laws_2|>
  mutate(Gender_Ministry_Presence = if_else(Gender_Ministry == "Gender Ministry", 1, 0))
prevalence_and_laws_3
```


```{r}
vaw_regression <- lm(PREVVIOLLIFETIME ~ Gender_Ministry_Presence + LAWDOMVIOL, data = prevalence_and_laws_3)
summary(vaw_regression)
# y = 22.680 + 1.818x1 + 8.747x2
```
```{r}
library(equatiomatic)
extract_eq(vaw_regression)

extract_eq(vaw_regression, use_coefs = TRUE)
```



## Discussion and Limitations
<!-- a brief (one paragraph) concluding section that summarizes your results, assesses the extent to which you find support for your hypothesis, describes limitations of your analysis and threats to inference, and states how your analysis could be improved (e.g., improved data that would be useful to collect) -->
<!-- A conclusion section, including a recap of results, discussion of limitations of the study, and how to improve the study in the future (a paragraph) -->

Within this project, I did not find support for my hypothesis.  Rates of violence were actually slightly higher for countries with a gender ministry, although this was not statistically significant.  The only measured relationship with a strong statistical significance was between (explain here).  However, even though I did not find support for my hypothesis, I still believe it to be true.  This project was severely limited by constraints of the data, which included only (insert amount here) countries reported, in comparison to the 193 that are typically considered to be the gold standard of country research, like in the United Nations.  Moreover, while each country had a measure of which laws that were reported based on the laws in their nation, since violence against women was self-reported, countries that scored high on laws which discriminated against women tended to be the ones that did not submit measures. Thus, the only countries that did submit rates could have been the ones with more transparent systems.  On a similar note, countries with a gender ministry or higher rates of gender equality still have violence against women, as it is a problem in every nation, but may have a better reporting system which empowers women to report and makes them feel safe when they do, which could result in higher reported rates of violence, even if violence overall may be lower.  Generally, violence against women is extremely underreported, and this is likely worse in countries where women are less safe. Additionally, in the future, I would want to improve my data by breaking down the gender ministry variable and creating an objective spectrum, with no gender ministry as zero, an actual gender ministry as one, and a family, children, equality, or other similar ministry as somewhere in between.  Finally, other factors, including GDP of a country, could be included to control for the role of funding, which certainly impacts how successful a gender ministry can be.  If possible, I would also want to find the amount that the ministry is funded with, including both the currency amount and what percent of the GDP that was, to provide essential context on both the amount and what the amount is compared to, to show how much of a priority it is, how established it might be, and what is considered a reasonable amount to distribute on a country-by-country basis. Finally, I would like to find a more reliable method of data collection, including one which measured my variables of interest (i.e. laws and violence against women) at multiple points in time, which I could then use to incorporate time of the gender ministries, including when they were founded, which is a really important area that I was unable to incorporate, due to a lack of variety in years reported from the OECD.  

## Link
<!-- there must be a link to the GitHub repo that contains the source code to load, clean, and analyze the data and produce the article -->
